Efficient and Truthful Forecasting Competitions

### 2021-01-12 paper/draft2.tex
new draft

\section{Introduction}
- SHELF\\
- GNOME\\
- ???

keywords: event, forecast, voting

EFV

@@

Let $F_{i, k}$ be the indicator function for forecaster $i$ winning the lottery for event $k$. Then, $F_i = \sum_k F_{i, k}$ is the random variable that is the number of events forecaster $i$ wins. The proof of Theorem 8 \cite{witkowski2018incentive} utilized the Hoeffding bound applied to $F_i$, which assumes they have high variance. However, this is not the case, and utilizing that fact gives an improvement in the lower bound. 
\begin{theorem}[Bernstein's Inequality]
  \label{bernstein}
  Given random variables $X_i$ for $1 \leq i \leq n$ such that $0 \leq X_i \leq 1$ almost surely, let $S = \sum_i X_i$. Then,
  \[ P\left(|S - \E[S]| < t\right) < 2 \: e^{\frac{-t^2 }{2\left(\Var(S) + \frac{t}{3}\right)} } \]
\end{theorem}
\begin{theorem}
  \label{elf_bound}
    $m_\delta(\elf, n, \epsilon) \leq \frac{5(n-1)}{\epsilon^2}\ln\left(\frac{4(n-1)}{\delta}\right)$
    % For an accuracy gap $\epsilon$, ELF correctly chooses the most accurate forecaster with probability at least $1 - 4 (n-1) e^{-\frac{m \epsilon}{6(n - 1)}}$.
\end{theorem}
\emph{Proof:} Note that for any single event $k$, every forecaster wins with probability at best $\frac{2}{n}$, and at worst $\frac{1}{n} - \frac{1}{n(n-1)}$ (achieved when $y_i = 1$, $y_{j\neq i} = 0$, and $x_i = 1$). Since $F_{i, k}$ is just a binary random variable, it will have highest variance when it's expectation is as close to $\frac{1}{2}$ as possible. For nontrivially small $n$, this will mean its expectation is $\frac{2}{n}$ (or equivalently $P(F_{i, k} = 1) = \frac{2}{n}$), so its variance will be at most $\frac{2(n-2)}{n^2}$. So we have $\Var(F_{i, k}) \leq \frac{2(n-2)}{n^2} < \frac{2}{n}$.

@@

\section{SHELF}

For this mechanism, we let the binary random variable $X_{i, k}$ be 1 with probability $R_q(y_{i, k}, x_i)$, and 0 otherwise. Then, for each event $k$ we assign forecaster $i$ the score $f_{i, k} = X_{i, k}$ obtained by sampling this variable. Finally, we assign their final score by summing these over all events, $F_i = \sum_{j=1}^m f_{i, j}$, then chose the forecaster with the highest score $\shelf(y, x) = \argmax_i F_i$, breaking ties uniformly at random.

\subsection{Truthfulness}
Note: The following analysis is inspired by and follows the work done in \citet{witkowski2018incentive}.

(This is wrong because we can't just look at their probability of winning a point. We need to also factor in the tiebreaker weighting some wins differently)

For a single event with true probability $\theta$, every forecaster will either win 1 or 0 events. The probability that their final score is 1 will just be their expected score for the single event. Therefore, to maximize their final score, each forecaster must maximize their expected score for the single event. Their score is just $\E_{x \sim \theta}(R_q(y_i, x)) = \E_{x \sim \theta}(1 - (y_i - x)^2)$. This is maximized when $y_i = \theta$, so the mechanism is truthful. We can also see this directly by noting that $R_q$ is a strictly proper scoring rule, so its expectation is maximized when $y_i = \theta$.

\begin{theorem}
  \label{shelf_truthful}
    $\shelf$ is truthful for $m \geq 1$ events.
\end{theorem}
\emph{Proof:} For each forecaster $j \in [1, n]$ fix their report.

Suppose that for some forecaster $i$, $i$'s report on some subset of events, where $k \in [1, m]$ denotes the size of this subset, does not report their true belief for any of the $k$ events.\\
Let $k^\prime$ be one of those events.

We will now show that $i$ will maximize their probability of being chosen as the best forecaster in the SHELF mechanism iff they alter their $k$ reports to report their true belief on each of the $k$ events. A guiding intuition here would be as follows: The higher the expected score a forecaster has on a given event, the higher the probability that SHELF awards them a point on this specific event, and the only way to maximize one's probability of being selected as the best forecaster is to increase the number of points they are awarded. Therefore, for a single independent event, increasing one's expected score will either increase the probability of being selected as the best forecaster or leave it unchanged, and it does so without altering results of any other event. In this way, we will show that being truthful for a singular event will always increase one's probability of being selected as the best forecaster, because, using a quadratic scoring rule, a forecaster's expected score is maximized when reporting their true belief. We can then repeat this process for each untruthful report to maximize the probability of being selected as the best forecaster.\\\\
Formally, we will analyze three cases relating to the context provided by the $m - 1$ events other than $k^\prime$.

Let $V(i)$ denote the number of points forecaster $i$ has accrued.\\
Let $f_{i,k}$ denote the probability forecaster $i$ scores a point on event $k$.\\

\begin{enumerate}
    \item Some forecaster $j$ scores at least two more points than $i$, OR $i$ scores at least two more points than any other forecaster.\\
    $$\exists j \in [1,m] \text{ s.t. } j \ne i \text{ and } V(j) - V(i) \geq 2$$ 
    $$OR$$
    $$\forall j \in [1,m] \text{ s.t. } j \ne i; V(i) - V(j) \geq 2$$
    For this case, $i$'s probability of being selected as the best forecaster is either exactly $1$ or $0$. It is also unaffected by being truthful or not in $k^\prime$, because gaining, or failing to gain, $1$ point in this context is not enough to alter the outcome as only the forecaster with the most number of points is chosen as the best forecaster with ties broken uniformly.

    \item Some forecaster $j$ scores exactly one more point than $i$, AND no other forecasters score strictly more points than $i$.\\
    $$\exists j \in [1,m] \text{ s.t. } j \neq i \text{ and } V(j) - V(i) = 1$$
    $$AND$$
    $$\forall l \in [1,m] \text{ s.t. } l \neq i \text{ and } l \neq j; V(i) - V(l) \geq 0$$
    For this case, $i$ must seek to maximize $f_{i,k^\prime}$ to have any non-zero probability of being selected as the best forecaster under the SHELF mechanism. 
    In the SHELF mechanism, the maximum payout a forecaster can receive for a single event is $1$, and that point is awarded to them with probability exactly equal to their quadratic score. Thus, we know a forecaster's expected value for a singular event is exactly their quadratic score. Since the quadratic scoring rule is proper, we know that the expectation of a forecaster's quadratic score is maximized when the forecaster is reporting their true beliefs. As a result, we can conclude that the only way for $i$ to maximize their probability of receiving a point in $k^\prime$ is to report their true beliefs. By reporting their true beliefs, which maximizes the probability that they score another point, $i$ also maximizes the probability of being selected as the best forecaster which, in this context, occurs when only $i$ scores a point and the tie between $j$ and $i$ is broken uniformly via the rules of SHELF.

    \item No forecasters score more points than $i$, but there is at least one forecaster $j$ that scores the same number of points as $i$ or scores exactly one less point than $y_i$.
    $$\exists j \in [1,m] \text{ s.t. } j \ne i \text{ and } V(i) - V(j) \in \{0,1\}$$
    In any given configuration, a forecaster's probability of being selected as the best forecaster is maximized when the minimal number of other forecaster have as many points as them. In the SHELF mechanism, any forecaster's probability of scoring a point on a given event is independent of all others. Therefore, there is no action that can be taken to minimize the probability of any competitor's probability of being selected as the best forecaster except strictly increasing one's own point total. In this final case, the only way for $i$ to increase their probability of being chosen as the best forecaster would be to increase the probability of scoring a point in $k^\prime$. The same logic as case 2 holds here that $i$ would then prefer to be truthful and maximize their expected score in $k^\prime$ to increase their probability of being selected as the best forecaster.
\end{enumerate}

Since any configuration of the $m-1$ events falls into one of these three cases, we have shown that any forecaster is weakly incentivized to report truthfully.
To show that any forecaster is strongly incentivized to report truthfully, we simply show that $i$ believes cases 2 or 3 will happen with positive probability. Since $y_{j,k} = p_{j,k} \in (0,1) \forall j \forall k$, we know that every forecaster has some uncertainty on each event. As a result, we can conclude that $f_{j,k} > 0 \forall j \forall k$. We can think of all possible outcomes as the set of all possible binary strings of length $n$, and since $f_{j,k} > 0 \forall j \forall k$ there exists at least one sequence of outcomes strings picked from our set that constitutes cases 2 or 3. As a result, there is a positive probability that cases 2 or 3 occur, and thus forecasters are strongly incentivized to report truthfully.

\subsection{Accuracy}

% From Proposition 2 of \citet{witkowski2018incentive}, we have that $E_{x \sim \theta}(R_q(y, x)) = R_q(y, \theta) + \theta^2 - \theta$. Additionally, for a single event, a forecaster's accuracy $a_i$ is by definition just $R_q(y, \theta)$. So, $E_{x \sim \theta}(R_q(y, x)) = a_i + \theta^2 - \theta$. For a single event with probability $\theta$, forecaster $i$'s expected score is $\E_{x \sim \theta} f_i = E_{x \sim \theta}(R_q(y_i, x)) = a_i + \theta^2 - \theta$.

% \subsection{Truthfulness for a single event}
% (section 4.3, proposition 5).  For forecaster $i$ to be selected, they either need to have $f_i = 1$, and then be chosen uniformly at random from all others who scored 1, or everyone must score 0 and they will be chosen uniformly at random. In the first case, the expected number of other forecasters who also score 1 is $\sum_{i' \neq i} \E_{x \sim \theta} f_{i'} = \sum_{i' \neq i} (a_{i'} + \theta^2 - \theta) = (n-1) (\theta^2 - \theta) \sum_{i' \neq i} a_{i'}$. Then, the probability that forecaster $i$ is chosen is 
% \[
%   P(f_i = 1) \frac{1}{1 + (n-1) (\theta^2 - \theta) \sum_{i' \neq i} a_{i'}} = \frac{a_i +\theta^2 - \theta}{1 + (n-1) (\theta^2 - \theta) \sum_{i' \neq i} a_{i'}}
% \]
% In the other case, the probability $i$ wins is just 
% \[
%   \frac{1}{n} \prod_{i} P(f_i = 0) = \frac{\prod_i (1 - a_i -\theta^2 + \theta)}{n}
% \]

% There's probably a better way to do this... Actually coming back to this I'm not even sure that the first step is allowed (using expected number of winners directly instead of summing over cases)

% \subsection{Truthfulness for multiple events}
\begin{theorem}
  \label{shelf_bound}
    $m_\delta(\shelf, n, \epsilon) \leq \frac{2}{\epsilon^2} \ln \frac{4(n-1)}{\delta}$
\end{theorem}
\emph{Proof:} This proof is a slight modification of the proof of Theorem 8 in \cite{witkowski2018incentive}.

We first bound the difference in the expected number of wins of $i^*$ and $j$, for any $j \neq i^*$. 
\begin{align*}
  \E \left( \sum_{k=1}^m f_{i^*, k}\right) - \E \left( \sum_{k=1}^m f_{j, k}\right)
  &= \left( \sum_{k=1}^m \E f_{i^*, k}\right) - \left( \sum_{k=1}^m \E f_{j, k}\right)\\
  &= \sum_{k=1}^m \left(\E (f_{i^*, k}) - \E (f_{j, k})\right)\\
  &= \sum_{k=1}^m \left((a_{i^*} + \theta_k^2 - \theta_k) - (a_j + \theta_k^2 - \theta_k)\right)\\
  &= \sum_{k=1}^m \left(a_{i^*} - a_j\right)\\
  &\geq \sum_{k=1}^m \epsilon\\
  &=m \epsilon
\end{align*}
If $F_j \geq F_{i^*}$, then either $\E(F_{i^*}) - F_{i^*} \geq \frac{m \epsilon}{2}$ or $F_j - \E(F_j) \geq \frac{m \epsilon}{2}$. By Hoeffding's, we can upper bound the probability of each of these events by $2 e^{-\frac{2 (\frac{m \epsilon}{2})^2}{m}} = 2 e^{\frac{- m \epsilon^2}{2}}$. Using the union bound, the probability that either of those events occurs is twice this, or $4 e^{\frac{- m \epsilon^2}{2}}$.

Using the union bound again over all $j \neq i$, we have 
\begin{align*}
  P(\shelf(y, x) = i) &= P(F_{i^*} > F_j \quad \forall j \neq i^*) \\
                      &\geq 1 - \sum_{j\neq i^*} P(F_{i^*} \leq F_j) \\
                      &= 1 - 4 (n-1) e^{\frac{- m \epsilon^2}{2}}
\end{align*}

Letting this quantity be $1 - \delta$ and solving for $m$, we see that this mechanism correctly chooses forecaster $i$ with probability $1 - \delta$ when $m$ is large enough to satisfy the inequality. Therefore
\[ m_\delta(\shelf, n, \epsilon) \leq \frac{2}{\epsilon^2} \ln \frac{4(n-1)}{\delta} \]
\hfill \QED

Comparing this to the result of Theorem \ref{elf_bound}, we see that $\shelf$ is much more efficient than $\elf$. We can also directly apply the proof of Theorem \ref{shelf_bound} to $M_s$, since the mean and variance of $F_i$ is the same in both cases. Therefore, $m_\delta(\shelf, n, \epsilon) = m_\delta(M_s, n, \epsilon)$.

\newpage
\section{Raf notes}

Let's study the one-round version of ELF + SHELF + noise.
And actually, we won't really need to use SHELF per se.
Let's actually define
$M_{f,\sigma}(y,x) = \max_{i\in[n]} f(y,x)_i + X_i$ where the $X_i \sim N(0,\sigma^2)$ are indpendent, and $f$ is any function with $f(y,x) \in [0,m]^n$ where if you change only one entry of $y$ by $c$, then $f$ changes by at most $c^2$.
(For example, $f$ could be the sum of scores.)
Let $M_\alpha$ be the mechanism that runs $M_{f,\sigma}$ with probability $\alpha$ and $\elf$ with probability $1-\alpha$.

In what follows, when I omit $x$, I'm implicitly taking the randomness over $x$ into account (from $\theta$).
Assume $y$ is truthful for forecaster $i$ and $y'$ is any vector that differs only in $i$'s forecast on a particular round.
And actually let's focus on the case $m=1$, so there is only one round anyway.
From a direct calculation (see also a Prop.\ in \cite{witkowski2018incentive}), we have, from forecaster $i$'s point of view,
\begin{align}
  \Pr[\elf(y)=i] - \Pr[\elf(y')=i] = \tfrac 1 n (y_i - y'_i)^2~,
\end{align}
since forecaster $i$ believes $y_i = \theta_1$.

By differential privacy (I have not checked this carefully!), and the fact that $\exp(c) \leq 1+2c$ for $c\in[0,1]$, we have, for each $x$,
\begin{align}
  \Pr[M_{f,\sigma}(y',x)=i]
  & \leq \exp\left( \frac{(y_i-y'_i)^2}{2\sigma^2} \right) \Pr[M_{f,\sigma}(y,x)=i]
  \\
  & \leq \left(1 + \frac{(y_i-y'_i)^2}{\sigma^2}\right) \Pr[M_{f,\sigma}(y,x)=i]
  \\
  & \leq \Pr[M_{f,\sigma}(y,x)=i] + \frac{(y_i-y'_i)^2}{\sigma^2}~.
\end{align}

Combining, the probability of $i$ winning for $y$ and $y'$ can be compared as follows.
\begin{align}
  &\Pr[M_\alpha(y)=i]-\Pr[M_\alpha(y')=i]\\
  &=\alpha\left(\Pr[M_{f,\sigma}(y)=i] - \Pr[M_{f,\sigma}(y')=i]\right)\\
  &\qquad + (1-\alpha)\left(\Pr[\elf(y)=i] - \Pr[\elf(y')=i]\right)~.
  \\
  &\geq -\alpha \frac{(y_i-y'_i)^2}{\sigma^2} + (1-\alpha) \frac{(y_i-y'_i)^2}{n}
  \\
  &\geq \left((1-\alpha) \frac{1}{n} -\alpha \frac{1}{\sigma^2}\right) (y_i-y'_i)^2~.
\end{align}
Thus, setting $\sigma^2 = n \tfrac \alpha {1-\alpha}$ gives truthfulness, e.g.\ $\sigma^2=n$ for $\alpha=\tfrac 1 2$.
Which is maybe not what we were hoping for, but it's something.

@@

%\begin{thebibliography}{99}
%\bibitem{elf} Witkowski, Freeman  \emph{Incentive-Compatible Forecasting Competitions} \url{https://las.inf.ethz.ch/files/witkowski-etal-aaai2018.pdf}.
%
%\bibitem{bnb} Raab, Steger  \emph{“Balls into Bins” — A Simple and Tight Analysis} \url{http://www.dblab.ntua.gr/~gtsat/collection/scheduling/Balls%20into%20Bins%20A%20Simple%20and%20Tight%20Anal.pdf}.
%\end{thebibliography}

>>>>

\section{Introduction}

@@

Let $F_{i, k}$ be the indicator function for forecaster $i$ winning the lottery for event $k$. Then, $F_i = \sum_k F_{i, k}$ is the random variable that is the number of events forecaster $i$ wins. The proof of Theorem 8 \cite{witkowski2018incentive} utilized the Hoeffding bound applied to $F_i$, which assumes they have high variance. However, this is not the case, and utilizing that fact gives an improvement in the lower bound. 
\begin{theorem}[Bernstein's Inequality]
  \label{bernstein}
  Given random variables $X_i$ for $1 \leq i \leq n$ such that $0 \leq X_i \leq 1$ almost surely, let $S = \sum_i X_i$. Then,
  \[ P\left(|S - \E[S]| < t\right) < 2 \: e^{\frac{-t^2 }{2\left(\Var(S) + \frac{t}{3}\right)} } \]
\end{theorem}
\begin{theorem}
  \label{elf_bound}
    $m_\delta(\elf, n, \epsilon) \leq \frac{5(n-1)}{\epsilon^2}\ln\left(\frac{4(n-1)}{\delta}\right)$.
    % For an accuracy gap $\epsilon$, ELF correctly chooses the most accurate forecaster with probability at least $1 - 4 (n-1) e^{-\frac{m \epsilon}{6(n - 1)}}$.
\end{theorem}
\emph{Proof:} Note that for any single event $k$, every forecaster wins with probability at best $\frac{2}{n}$, and at worst $\frac{1}{n} - \frac{1}{n(n-1)}$ (achieved when $y_i = 1$, $y_{j\neq i} = 0$, and $x_i = 1$). Since $F_{i, k}$ is just a binary random variable, it will have highest variance when it's expectation is as close to $\frac{1}{2}$ as possible. For nontrivially small $n$, this will mean its expectation is $\frac{2}{n}$ (or equivalently $P(F_{i, k} = 1) = \frac{2}{n}$), so its variance will be at most $\frac{2(n-2)}{n^2}$. So we have $\Var(F_{i, k}) \leq \frac{2(n-2)}{n^2} < \frac{2}{n}$.

@@

\subsection{Other point-per-round mechanisms}
We can extend Theorem \ref{elf_lower_bound} to apply to any normal, anonymous mechanism that assigns a single point per event to 1 forecaster, and then picks the forecaster with the highest score. We can treat such a forecaster selection mechanism as a budget-balanced wagering mechanism for each event, where each forecaster has a shared, fixed wager of $\frac{1}{n}$. Since we want the mechanism to by symmetric, we can assume it satisfies anonymity as described in [wagering mechanism paper].

(TODO: justify normality).

Then, by Lemma 4 in [wagering mechanism paper], we have that forecaster $i$'s payout for an event will be 
\[
\frac{1}{n} + f(y_i, x) - \frac{1}{n-1} \sum_{j \neq i} f(y_j, x)
\]
where $f$ is a proper scoring rule whose values are contained in an interval of size $\frac{1}{n}$.

@@
